---
title: "Tutorial8"
author: "Yangkai Hong"
date: "25/09/2018"
output: html_document
---
##The library “mlbench” contains a set of benchmark datasets for learning problems. Install this package and load the Ionosphere dataset from this package by typing “data(Ionosphere)”.
```{r}
library(mlbench)
data("Ionosphere")
Ionosphere <- Ionosphere[,-1]
Ionosphere <- Ionosphere[,-1]
```
###(1) Partition the dataset into training and test sets (60% and 40%).
```{r}
set.seed(1)
inTrain <- createDataPartition(Ionosphere$Class, p = 0.6)[[1]]
IonosphereTrain <- Ionosphere[inTrain,]
IonosphereTest <- Ionosphere[-inTrain,]
```
###(2) Using the absolute fold change to filter the data for classification. Plot the number of features included against the classifier performance of kNN and logistic regression on test set.
```{r}
IonosphereTrain.byClass <- split(IonosphereTrain[,-33],IonosphereTrain$Class)
feature.mean.byClass <- sapply(IonosphereTrain.byClass,colMeans)
# add abs inside log2
feature.foldChange <- abs(log2(abs(feature.mean.byClass[,1] / feature.mean.byClass[,2]))) 
feature.sorted <- sort(feature.foldChange,decreasing = TRUE)
filtered.features1 <- names(feature.sorted)
features <- c()
accuracies <- c()
for(i in 1:length(filtered.features1)){
  features <- c(features,filtered.features1[i])
  features
  knn.filtered <- knn(train = IonosphereTrain[,features],test = IonosphereTest[,features],cl = IonosphereTrain$Class,k=5,prob = TRUE)
  knn.filtered
}
```
###(3) Repeat the above but using t-statistics to rank and filter the features this time.
```{r}
IonosphereTrain.byClass <- split(IonosphereTrain[,-33],IonosphereTrain$Class)
feature.pvalues <- c()
for(i in 1:(ncol(IonosphereTrain)-1)){
  feature.pvalues <- c(feature.pvalues,t.test(IonosphereTrain.byClass[[1]][,i],IonosphereTrain.byClass[[2]][,i])$p.value)
}
names(feature.pvalues) <- colnames(IonosphereTrain[,-33])
filtered.features2 <- names(sort(feature.pvalues))[1:10]
knn.filtered <- knn(IonosphereTrain[,filtered.features2],IonosphereTest[,filtered.features2],IonosphereTrain$Class,k=5,prob = TRUE)
table(knn.filtered,IonosphereTest$Class)
```
###(4) Compare the results from using absolute fold change and t-statistics. How many features should be included according to each method? Does your initial partition of training and test sets have anything to do with the selected features?
```{r}

```
